---
output: html_document
---

```{r global_options, include = FALSE}
#### Load any necessary libraries here. I've already loaded the basics
library(tidyverse)
library(conflicted)

filter <- dplyr::filter

knitr::opts_chunk$set(echo = TRUE, fig.align ="center", 
                      fig.height = 4.5, fig.width = 4.5)
```

## Exploratory Data Analysis

## Introduction

Here's where you'll explain your data. Where is it from and what's a little bit of the background. Then you need to explain the columns (variables) in the dataset:
  
- **column1:** description
- **column2:** description
- **column3:** description
- etc.

<br><br>

## Data Preparation

Describe here what you did to clean the data.

```{r}
### Use this chunk to read in the data and clean it
```

<br><br>

## Questions

For each question, make a plot illustrating the question, use a statistical to answer the question, and describe your conclusions.

#### Question #: WRITE YOUR QUESTION HERE

DESCRIBE YOUR RESULTS HERE

```{r}
### use this chunk to make a plot illustrating your question
```

```{r}
### use this chunk to conduct a statistical test to answer your question
```

**Rinse and repeat for another 9 questions**

## References
1. Reference should be for your dataset(s), but if you looked up more information that you included in the introduction, please reference it here as well.



<br>


