
---
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(conflicted)

# resolve conflicts
filter <- dplyr::filter

knitr::opts_chunk$set(echo = TRUE)
```

## Assignment 2021-07-08: Plotting with `ggplot2`

We'll be using data from the `gapminder` package for today's assignment. The `gapminder` dataset has the following columns (if the name defines the column I didn't put a longer description): 

- `country`
- `continent` 
- `year` 
- `ifeExp` = the life expectancy of the population in years
- `pop` = the population of the country
- `gdpPercap` = gross domestic product (GDP) per capita in inflation-adjusted US dollars 

You need to install and load a package to get the data, so make sure you pay attention to the set up steps and run every chunk. 

### Set Up

Install the package with the data

```{r}
# install.packages("gapminder")
```

Load the library.

```{r}
library(gapminder)
```

Look at the data. 

```{r}
gapminder
```

<br>

### Play with the Fixed Aesthetics

We're going to play around with the fixed aesthetics of the same plot

---

1. Change the transparency of the points

```{r}
gapminder %>%
  filter(gdpPercap < 50000) %>%
  
ggplot(aes(x = gdpPercap, y = lifeExp)) +
  geom_point(aes(color = continent)) +
  geom_smooth(se = F)
```

2. Make the points and/or the line a different color! A pdf of R color names is available at the link <http://www.stat.columbia.edu/~tzheng/files/Rcolor.pdf

```{r}
gapminder %>%
  filter(gdpPercap < 50000) %>%
  
ggplot(aes(x = gdpPercap, y = lifeExp)) +
  geom_point(aes(color = continent)) +
  geom_smooth(se = F)
```

3. You don't have to have just a boring straight lines or normal circles! Change the line to a different type. Options at the link: <http://sape.inf.usi.ch/quick-reference/ggplot2/linetype>. Change the points to something other than circles too. <https://ggplot2.tidyverse.org/articles/ggplot2-specs_files/figure-html/unnamed-chunk-8-1.png>

```{r}
gapminder %>%
  filter(gdpPercap < 50000) %>%
  
ggplot(aes(x = gdpPercap, y = lifeExp)) +
  geom_point(aes(color = continent)) +
  geom_smooth(se = F)
```

4. Change the size of something on the plot! The `size` argument accepts numeric values.

```{r}
gapminder %>%
  filter(gdpPercap < 50000) %>%
  
ggplot(aes(x = gdpPercap, y = lifeExp)) +
  geom_point(aes(color = continent)) +
  geom_smooth(se = F)
```

5. Change the theme of the plot to something else. Options at the link: <https://ggplot2.tidyverse.org/reference/ggtheme.html>

```{r}
gapminder %>%
  filter(gdpPercap < 50000) %>%
  
ggplot(aes(x = gdpPercap, y = lifeExp)) +
  geom_point(aes(color = continent)) +
  geom_smooth(se = F)
```

6. Facet the plot by some feature

```{r}
gapminder %>%
  filter(gdpPercap < 50000) %>%
  
ggplot(aes(x = gdpPercap, y = lifeExp)) +
  geom_point(aes(color = continent)) +
  geom_smooth(se = F)
```

7. Change the x- and y- axis labels of the plot to something better

```{r}
gapminder %>%
  filter(gdpPercap < 50000) %>%
  
ggplot(aes(x = gdpPercap, y = lifeExp)) +
  geom_point(aes(color = continent)) +
  geom_smooth(se = F)
```

### Same Data, Different Shapes, Different Geometry

This section will explore how you can plot data with different geometries

---

For questions 8 and 9, plot the gapminder dataset with `x = continent` and `y = lifeExp`. Use any geom you would like.

8. `x = continent` and `y = lifeExp` Plot 1

```{r}
ggplot(gapminder, aes(x = continent, y = lifeExp)) +
  # geom of your choice goes here
```

9. `x = continent` and `y = lifeExp` Plot 2

```{r}
ggplot(gapminder, aes(x = continent, y = lifeExp)) +
  # geom of your choice goes here
```

---

For questions 10-12, pick one variable and plot it using a geom of your choice

10. Pick one variable and plot it

```{r}

```

11. Pick one variable and plot it

```{r}

```

12. Pick one variable and plot it

```{r}

```

<br>

### Answer Question With Plots

For this section, make an appropriate plot to answer the question.

---

13. What is the trend in life expectancy over time? Is it increasing or decreasing?

```{r}

```

14. As life expectancy changes, what happens to GDP per capita?

```{r}

```

15. Does the relationship between life expectancy and GDP per capita look the same for all continents?

```{r}

```

16. Is the distribution of life expectancy the same over all the continents?

```{r}

```

17. Is the relationship between life expectancy and GDP per capita the same in the first year in the `gapminder` dataset vs the last year?

```{r}

```

18. Which country in Oceania had the greatest increase in life expectancy over time?

```{r}

```

19. What's the difference in population between the countries of Oceania in the 21st century?

```{r}

```

20. For countries starting with 'B', how has their GDP per capita changed over time?

```{r}

```



<br><br>


