---
title: ""
output: html_document
---

```{r setup, include=FALSE}
library(magrittr)
library(tidyverse)
library(vroom)

library(ggupset)

knitr::opts_chunk$set(echo = TRUE)
```

# 2021-07-21: RNA-Seq Clustering and Differential Expression Testing

Please email me your Rmarkdown document at <kkeith@coriell.org> with your **LAST NAME** in the filename by 9AM Tuesday 7/27.

---

<br>

---

## Lecture Questions

For the multiple choice questions, please select your answer in **bold**.

---

1. What is the correct interpretation of an MA plot?
  a. As a gene moves farther away from the horizontal line running through 0 on the y-axis, the greater the change in gene expression between conditions.
  b. The closer a gene is to the horizontal line running through 0 on the y-axis, the greater the change in gene expression between conditions.
  c. Only genes that have low mean expression can be differentially expressed.
  d. The MA plot shows the statistical significance of all genes tested.

---

2. What is the correct interpretation of a volcano plot?
  a. As you move farther away from 0 on the x-axis, the difference in expression between conditions decreases and as you move up the y-axis the statistical significance of the difference increases, so the points at the top middle of the plot are the most significantly differentially expressed genes.
  b. As you move farther away from 0 on the x-axis, the difference in expression between conditions decreases and as you move up the y-axis the statistical significance of the difference decreases, so the points at the center bottom of the plot are the most significantly differentially expressed genes.
  c.  As you move farther away from 0 on the x-axis, the difference in expression between conditions increases and as you move up the y-axis the statistical significance of the difference increases, so the points at the top right and left of the plot are the most significantly differentially expressed genes.
  d. As you move farther away from 0 on the x-axis, the difference in expression between conditions increases and as you move up the y-axis the statistical significance of the difference decreases, so the points at the bottom right and left of the plot are the most significantly differentially expressed genes.

---

3. What is the correct interpretation of an UpSet plot?
  a. Having bars of all equal heights means there's a large overlap between sets.
  b. Any tall bar means that there's a large overlap between sets.
  c. Having tall bars at the single dots means there's a large overlap between sets.
  d. The points and lines at the bottom show the intersection between the sets, so if the bar is large where sets overlap, there is strong overlap between conditions.

---

<br>

---

## Exploring RNA-seq Results

### Read / Wrangle Data

#### Read Data In

Use the chunk below to read in your differential expression results from the last homework.

```{r}

```

### Wrangle Data

Use the chunk below to get human readable gene names for your results

```{r}

```

Finally use the chunk below to combine your differential expression table, table of gene names, and to add useful columns for plotting to it. Make sure to add columns indicating:

- Whether or not the gene is significantly differentially expressed
- The direction the expression of the gene changes relative to control.
- The -log10 of the q-value

```{r}

```

<br>

### Visualize Results

#### MA Plot

Use the chunk below to plot an MA plot

```{r}

```

#### Volcano Plot

Use the chunk below to plot a volcano plot

```{r}

```

#### UpSet Plot

Use the chunk below to plot an UpSet plot

```{r}

```




<br>



