---
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(conflicted)

### resolve conflicts
filter <- dplyr::filter

knitr::opts_chunk$set(echo = TRUE, fig.width = 6, fig.height = 4)
```

# Tidy Data Wrangling with `dplyr`

`dplyr` is a grammar of data manipulation, providing a consistent set of rationally named functions to help you solve the most common data manipulation challenges.

<br>

## Subsetting Data in Base `R`

### Tables and Examing Data

```{r}
iris
```

`head()` and `tail()`

```{r}
# look at the first 6 rows
head(iris)

# look at the first 10 rows
head(iris, 10)

# look at the last 6 rows
tail(iris)
```

`dim()` prints the numbers of rows and columns in a table

```{r}
# rows, columns
dim(iris)
```

### Subsetting Tables with Base R

Square brackets `[]`

```{r}
iris[]
```

Dollar sign subsetting `$`

```{r}
iris$Sepal.Length
```

More advanced subsetting

```{r}
# drop the first column
iris[,-1]

# select rows 1 to 5 and columns 2 to 4
iris[1:5,2:4]

# list the exact rows and columns you want
iris[c(37, 49, 132, 12, 77), c(1, 3)]
```

### Logicals

These operators are how you compare values to each other. Keep them in mind for subsetting and filtering going forward.

- equals `==`
- not equal to `!=`
- greater than `>`
- greater than or equal to `>=`
- less than `<`
- less than or equal to `<=`
- and `&`
- or `|`

```{r}
# Are the numbers equal?
2 == 5

# Is the first number less than the second number?
2 < 5

# Is the first number greater than the second number?
2 > 5

# Are the numbers NOT equal
2 != 5
```

Subset tables with logicals

```{r}
# get all columns and all rows where sepal length is greater than or equal to 5
iris[iris$Sepal.Length >= 5,]

# get all columns and all rows species is NOT viriginica
iris[iris$Species != 'virginica',]
```

### Piping

The pipe operator `%>%` pipes output from one function to the next. You can use this to chain together as many command/functions as desired. For a simple example, instead of calling `head()` on flights, we could have piped flights to `head()`

```{r}
iris %>% head()
```

You can chain as many pipes as you want together and you can also pipe into `ggplot()`

```{r}
iris %>% filter(Species == 'setosa') %>% select(Sepal.Length, Sepal.Width)
```

### Assignment

The assignment operator, `->` saves the current output as an object (aka a data structure) to whatever name you pick. You can save data of any format this way.

```{r}
iris %>% filter(Species == 'setosa') %>% select(Sepal.Length, Sepal.Width) -> new_iris

new_iris
```

<br>

## Data Manipulation with dplyr

### Setup

Install packages

```{r}
# uncomment the packages and install
# install.packages('tidyverse')
# install.packages('conflicted')

# if the packages aren't loaded run the topmost chunk in this document
```

### dplyr functions

#### `arrange()`

The `arrange()` function allows you to sort the data by some criteria, usually another column in the data. 

```{r}
### Before arranging the data, let's remember what it looks like
iris

### arrange() takes, at a minimum, a single column as an argument. It will sort
### the table by that column in order from least to greatest
iris %>%
  arrange(Petal.Width)

### You can sort the data by as many columns as you like, although the changes
### in the order are smaller with each additional column
iris %>%
  arrange(Petal.Width, Petal.Length, Sepal.Length)

### What if you want to sort from largest to smallest? You can use the helper
### function desc() that changes the sort order from largest to smallest.
iris %>%
  arrange(desc(Petal.Width))
```

### `select()`

The `select()` function allows you to pick only the columns you want

```{r}
### For the select() function, you just give it the names of the columns you 
### want. For more complicated selection criteria, you probably need to use 
### select_if() or select_at() which are beyond the scope of this presentation.
iris %>%
  select(Species, Sepal.Length, Sepal.Width)

iris %>%
  select(Species:Sepal.Width) -> select_example
select_example

iris %>%
  select(species = Species, Sepal.Length, Sepal.Width)
```

### `filter()`

The function `filter()` allows you select a subset of the data according to some criteria. This could be by matching a variable like `sample == 's42'` or could be base on some numberic value like `Sepal.Width < 3 | Sepal.Width > 4`

```{r}
### filter takes that name of the column you want to filter on and some 
### logical condition. Here, we'll get only the rows of the table for samples
### with Sepal.Lengths greater than 4.5 cms.
iris %>%
  filter(Sepal.Length > 5.5, Species == 'setosa')

iris %>%
  filter(Sepal.Length > 5.5 | Species == 'setosa')

### You can filter by as many things as you want
iris %>%
  filter(Sepal.Length > 4.5, Species %in% c('virginica', 'versicolor'), 
         Sepal.Width < 3, Petal.Length > 5, Petal.Width > 1.5)
```

### `distinct()`

The function `distinct()` gives you all unique values of a given variable.

```{r}
### distinct() takes the name of the column you want the distinct values of
iris %>%
  distinct(Sepal.Width)

### You can give it as many columns as you want although it becomes less and
### less useful as you add columns
iris %>%
  distinct(Species, Sepal.Width)
```

### `mutate()`

The function `mutate()` adds an additional column onto the table.

```{r}
iris %>%
### mutate() needs the name of the new column equal to whatever will go in the
### new column. Frequently that will be a logical statement based on other
### data in the table, like in this example where we call any iris with a sepal
### length less than 5.8 cm small and all the other irises large
  mutate(flower_size = ifelse(Sepal.Length < 5.8, 'small', 'large'))

### You can make more than one column in the same mutate call. If you supply a
### single value, it will be used for all rows (recycled).
iris %>%
  mutate(flower_size = ifelse(Sepal.Length < 5.8, 'small', 'large'),
         second_column = 'tidyverse_webinar')

iris %>%
  mutate(flower_size = case_when(Sepal.Length < 3 ~ 'small',
                                 Sepal.Length >= 3 & Sepal.Length < 6 ~ 'medium',
                                 TRUE ~ 'large'))
```

### `summarize()`

Lets you reduce the data down to some summary statistics or other summary function

```{r}
### summarize() works similarly to mutate(), but instead of adding an 
### additional column with an entry for all rows, it will reduce to the number
### summaries you asked for. Like mutate(), summarize() needs the name of the
### summary column equal to the calculation you're making. Here, we'll make the
### summary column "average" which is equal to the mean of the petal lengths
iris %>%
  summarize(average = mean(Petal.Length))

### For most tidyverse functions, either the British or American spellings can
### be used. Frequently the British spelling comes up by default because Hadley
### Wickham is from New Zealand.
iris %>%
  summarise(average = mean(Petal.Length))

### summarize() is very powerful when used with group_by(). group_by() causes 
### all functions later in the pipeline to be applied over the specified group
iris %>%
  group_by(Species) %>%
  summarize(average = mean(Petal.Length))

### Don't forget to ungroup() because grouping will mess up your code later
iris %>%
  group_by(Species) %>%
  summarize(average = mean(Petal.Length)) %>%
  ungroup()

### You can calculate more than one summary at a time as long as they (for the
### moment) have a single thing as their output
iris %>%
  group_by(Species) %>%
  summarize(avg_pl = mean(Petal.Length),
            avg_pw = mean(Petal.Width),
            sd_pl = sd(Petal.Length),
            sd_pw = sd(Petal.Width)) %>%
  ungroup()
```

#### `count()`

The function `count()` is like a special case of `summarize()`, but all it does is count the number of distinct things in the column(s).

```{r}
### count() takes the name of the column you want
iris %>%
  count(Species)

### It's particularly powerful when combined with group_by(), like summarize() is
iris %>%
  mutate(flower_size = ifelse(Sepal.Length < 5.8, 'small', 'large')) %>%
  group_by(flower_size) %>%
  count(Species)

### You can also count() multiple things at once
iris %>%
  mutate(flower_size = ifelse(Sepal.Length < 5.8, 'small', 'large')) %>%
  count(Species, flower_size)
```



